---
title: "Chapter 6 Lab 3"
output: html_notebook
---

```{r}
library(ISLR)

attach(Hitters)

Hitters = na.omit(Hitters)
```

## 6.7.1 - Principal Components Regression

PCR can be performed using the `pcr()` function, part of the `pls` library.

```{r}
library(pls)

set.seed(647)

pcr.fit = pcr(Salary ~ .,
              data = Hitters,
              scale = TRUE, # Standardize predictors
              validation = 'CV') # ten-fold cv for each possible value of M (number of PCs used)
```

```{r}
summary(pcr.fit)
```

CV score is provided for PCs ranging from M = 0 (intercept only) to M = 19 (basically least square).

Note that `pcr()` reports the *root mean squared error*, so to get the usual MSE, we need to square it.

```{r}
validationplot(pcr.fit,
               val.type = 'MSEP') # Plot the MSE
```

We see that the lowest MSE occurs when M = 16, barely fewer than 19.

We can also see that, when M = 1, MSE is roughly the same. This suggests that a small number of components might suffice.

Now, let's performe PCR on the training data and evaluate its test set performance.

```{r}
set.seed(647)

train = sample(1:nrow(x), nrow(x) / 2) # 50/50 split

length(train)

train[1:20]

test = (-train)

y.test = y[test]

pcr.fit = pcr(
  Salary ~ .,
  data = Hitters,
  subset = train,
  scale = TRUE,
  validation = 'CV'
)

validationplot(pcr.fit, val.type = 'MSEP')
```

```{r}
summary(pcr.fit)
```

Lowest MSEP occurs at M = 17. Let's use M = 5.

```{r}
pcr.pred = predict(pcr.fit,
                   x[test,],
                   ncomp = 5)

mean((pcr.pred - y.test) ^ 2)
```

Finally, fit on the entire data set.

```{r}
pcr.fit = pcr(y ~ x, scale = TRUE, ncomp = 5)

summary(pcr.fit)
```

## 6.7.2 - Partial Least Squares

PLS is done using the `plsr()` function.

```{r}
set.seed(647)

pls.fit = plsr(
  Salary ~ .,
  data = Hitters,
  subset = train,
  scale = TRUE,
  validation = "CV"
)

summary(pls.fit)
```

```{r}
validationplot(pls.fit, val.type = "MSEP")
```

Lowest occurs with M = 12.

```{r}
pls.pred = predict(pls.fit, x[test,], ncomp = 12)

mean((pls.pred - y.test) ^ 2)
```

MSE is higher than using PCR.

Full data set, using M = 12.

```{r}
pls.fit = plsr(Salary ~ .,
               data = Hitters,
               scale = TRUE,
               ncomp = 12)

summary(pls.fit)
```

